---
title: "R-Replications and Simulations-2"
output: html_notebook
---
This program replicates the bounds indicated at the tops of Figures 2.7. from ``Using Instrumental Variables for Inference about Policy Relevant Treatment Parameters'' (2018, Econometrica) by Mogstad, Santos and Torgovitsky,

As a exercise I replicate the analysis using matrix algebra instead of packaged commands.

While the optimization runs under the specification provided, the results do not match.

Author: Zachary Tausanovitch 


```{r Setup, message=FALSE, warning=FALSE}

################################################################################
# Setup ########################################################################
################################################################################

set.seed(1162019)
setwd(
  'C:\\Users\\ztaus\\Dropbox\\Personal\\Programming\\R-Replications and Simulations-1'
)

# load in packages
package_list <- c("foreign", 
                  "MASS",
                  "matrixcalc",
                  "ggplot2",
                  "tidyverse",
                  "kableExtra",
                  "magick",
                  "gurobi")
invisible(lapply(package_list, library, character.only = TRUE))

```


```{r Functions}
################################################################################
# Functions ####################################################################
################################################################################

# build a function to calculate the values of the polynomials over the support using the relevant weights 
# Will define the weights below
bernstein_output <- function(lb,ub, fun,weightfn){
  Z<-0
  z0<- c(PZ_0 * weightfn(Z) * integrate(fun,0,p_0)[[1]], PZ_0 * weightfn(Z) * 
    integrate(fun,p_0,1)[[1]])
  Z<-1
  z1<- c(PZ_1 * weightfn(Z) * integrate(fun,0,p_1)[[1]], PZ_1 * weightfn(Z) * 
    integrate(fun,p_1,1)[[1]])
  Z<-2
  z2<- c(PZ_2 * weightfn(Z) * integrate(fun,0,p_2)[[1]], PZ_2 * weightfn(Z) * 
    integrate(fun,p_2,1)[[1]])
  
  answer <- z0 + z1 + z2
  return(answer)
}

# Define Berstein Polynomials as functions
m11<- function(u){(1-u)^2}
m12<- function(u){2*u*(1-u)}
m13<- function(u){(u)^2}
m01<- function(u){(1-u)^2}
m02<- function(u){2*u*(1-u)}
m03<- function(u){(u)^2}


# Another function for the OLS based optimization
bernstein_output_OLS <- function(lb,ub, fun,weightfn){
  D<-0
  d0<- c(PD_0 * weightfn(D) * integrate(fun,0,PD_0)[[1]], PD_0 * weightfn(D) * 
    integrate(fun,PD_0,1)[[1]])
  D<-1
  d1<- c(PD_1 * weightfn(D) * integrate(fun,0,PD_1)[[1]], PD_1 * weightfn(D) * 
    integrate(fun,PD_1,1)[[1]])
  
  answer <- d0 + d1
  return(answer)
}


# Define the Gurobi Equation Optimization Algorithm
# While there are built-in functions, this software
# is faster.
gurobi_fun <- function(A,obj,modsense,rhs,sense,lb,ub){

# follow the online instructions, 
  #with some small modifications for our implementation
model <- list()
model$A          <- A 
model$obj        <- matrix(obj, nrow=1, ncol=6, byrow = T)
model$modelsense <- modsense
model$rhs        <- rhs
model$sense     <- sense
# Set parameter bounds 
model$lb       <- lb
model$ub       <- ub

params <- list(OutputFlag=1)

result <- gurobi(model, params)

print('Solution:')
print(result$objval)
print(result$x)

rm(model, result, params)

}

```



```{r Data}
################################################################################
# Data #########################################################################
################################################################################

# As this is a simuuation, no data is required
```


################################################################################
 Analysis ######################################################################
################################################################################



```{r Question 5}
#Building the equation

# list what we know and can easily calculate
PD_0 <- .515
PD_1 <- .485

PZ_0 <- .5
PZ_1 <- .4
PZ_2 <- .1

p_0 <- .35
p_1 <- .6
p_2 <- .7

# Covariance calculation for OLS
print(((1-.485)^2*.485 + (0-.485)^2*.515))

# Calculate the weights (only the z and d are dependent)
IVweight <-  function(z) {(z-.6)/.089}
OLSweight <- function(d) {(d-.485)/.249775}

# Claculate and bind the outputs for the constraints
constraint_IV<- c(
bernstein_output(0,1,m01,weightfn=IVweight)[1],
bernstein_output(0,1,m02,weightfn=IVweight)[1],
bernstein_output(0,1,m03,weightfn=IVweight)[1],
bernstein_output(0,1,m01,weightfn=IVweight)[2],
bernstein_output(0,1,m02,weightfn=IVweight)[2],
bernstein_output(0,1,m03,weightfn=IVweight)[2])

constraint_OLS<- c(
bernstein_output_OLS(0,1,m01,weightfn=OLSweight)[1],
bernstein_output_OLS(0,1,m02,weightfn=OLSweight)[1],
bernstein_output_OLS(0,1,m03,weightfn=OLSweight)[1],
bernstein_output_OLS(0,1,m01,weightfn=OLSweight)[2],
bernstein_output_OLS(0,1,m02,weightfn=OLSweight)[2],
bernstein_output_OLS(0,1,m03,weightfn=OLSweight)[2])

# Calculate and bind the objective function
obj1 <- integrate(m11,lower = 0.35, upper = 0.9)[[1]]*(1/(.9-.35))
obj2 <- integrate(m12,lower = 0.35, upper = 0.9)[[1]]*(1/(.9-.35))
obj3 <- integrate(m13,lower = 0.35, upper = 0.9)[[1]]*(1/(.9-.35))
obj4 <- integrate(m01,lower = 0.35, upper = 0.9)[[1]]*(-1/(.9-.35))
obj5 <- integrate(m02,lower = 0.35, upper = 0.9)[[1]]*(-1/(.9-.35))
obj6 <- integrate(m03,lower = 0.35, upper = 0.9)[[1]]*(-1/(.9-.35))
objective <- cbind(obj1,obj2,obj3,obj4,obj5,obj6)
# Observe, how pretty. 
objective

# Insert the values we need 
EZ <- .6
cov_dz <- 0.089

# Use true theta to check that our constraints are correctly specified
true_theta <- c(0.75,0.5,0.25,0.6,0.4,0.3)
t(as.matrix(true_theta)) %*% as.matrix(constraint_IV)
t(as.matrix(true_theta)) %*% as.matrix(constraint_OLS)

# Specify all the different possible bounds
MTRconst  <- rbind(c(1,1,1,-1,-1,-1), c(1,1,1,-1,-1,-1))
MONOconst <- rbind(c(-1,1,0,0,0,0), c(0,-1,1,0,0,0), 
                    c(0,0,0,-1,1,0),c(0,0,0,0,-1,1))
CONCconst <- rbind(c(1,-2,1,0,0,0), c(0,0,0,1,-2,1))


# Collect the constraints we are using 
constraints_IV<- rbind(t(as.matrix(constraint_IV)),MTRconst,MONOconst,CONCconst)
constraints_IVOLS<- rbind(t(as.matrix(constraint_IV)),
  t(as.matrix(constraint_OLS)),MTRconst,MONOconst,CONCconst)
constraints_dec<- rbind(t(as.matrix(constraint_IV)),
  t(as.matrix(constraint_OLS)),MTRconst,MONOconst,CONCconst)

# relevant "true" values that will constrain us. 
beta_IV = 0.074
beta_OLS = 0.253

# collect bindings & direction they are bound 
rhsIV<- c(beta_IV,0,1, 0 ,0, 0, 0, 0, 0)
rhsIVOLS<- c(beta_IV,beta_OLS,0,1, 0 ,0, 0, 0, 0, 0)
senseIV       <- c('=','>=','<=', '<=', '<=', '<=','<=','<=','<=')
senseIVOLS     <- c('=','=','>=','<=', '<=', '<=', '<=','<=','<=','<=')

# use gurobi to optimize 
gurobi_fun(A=constraints_IV,obj=objective,modsense="max",rhs=rhsIV,
  sense=senseIV,lb=0,ub=1)
gurobi_fun(A=constraints_IV,obj=objective,modsense="min",rhs=rhsIV,
  sense=senseIV,lb=0,ub=1)

# Figure 2
print("IV")
gurobi_fun(A=constraints_IVOLS,obj=objective,modsense="max",rhs=rhsIVOLS,
  sense=senseIVOLS,lb=0,ub=1)

# Figure 3
print("IV & OLS")
gurobi_fun(A=constraints_IVOLS,obj=objective,modsense="min",rhs=rhsIVOLS,
  sense=senseIVOLS,lb=0,ub=1)


```